---
title: "Cross Validation"
author: "Kimberly Lopez"
date: "2024-11-14"
output: github_document
---

```{r}
library(tidyverse)
library(modelr)
library(mgcv)

library(SemiPar)

set.seed(1)
```

working with the Lidar dataset 

```{r}
data("lidar")

lidar_df = 
  lidar|>
  as_tibble()|>
  mutate( id= row_number())

```

Graphing the Lidar data: 
```{r}
lidar_df|>
  ggplot( aes(x= range, y = logratio))+ 
  geom_point()
```

**Is this model too complex to fit a linear relationship?** 

# Cross Validation 

We will compare 3 models, linear, smooth, wiggly 

Frist, contruct training and testing data 
```{r}

train_df = sample_frac(lidar_df, size = .8)

test_df = anti_join(lidar_df, train_df, by = "id")

```

Visualizing Training data: 
```{r}
ggplot(train_df, aes(x= range, y = logratio))+ 
  geom_point()+ 
  geom_point(data = test_df, color = "red")
```

Now we can start fitting one model: 
--> one model will be too simple or too complex and make incorrect predictions. 

Three models: 
```{r}
linear_mod= lm(logratio ~ range, data = train_df)
smooth_mod= gam(logratio ~ s(range), data = train_df)
wiggly_mod = gam(logratio ~ s(range, k = 30, sp= 10e-6), data = train_df)
```

Visualize the model fits: 

**Linear fit** 
```{r}
train_df|>
  add_predictions(linear_mod)|> 
  ggplot(aes(x=range, y = logratio))+ 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

- This is not complex enough since some of the datapoints are not linear 

**Wiggly model fit** 
```{r}
train_df|>
  add_predictions(wiggly_mod)|> 
  ggplot(aes(x=range, y = logratio))+ 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

- this over fits the training data on the model, so we cannot generalize this for a different dataset 

**Smooth model fit** 

- We use  `gam` when data is not linear 
```{r}
train_df|>
  add_predictions(smooth_mod)|> 
  ggplot(aes(x=range, y = logratio))+ 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

- this is better than linear since it captures the trends in the data, but not too flexible that will cause overfitting 

**Compare these numerically using RMSE** 

```{r}

rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

Smooth_mod has lower RMSE 

We want to see if this differece is consistent for random data testing 

# Repeat the Train/test split 

This creates 100 dataframes 
```{r}
cv_df = 
  crossv_mc( lidar_df, 100)|>
  mutate(
    train= map(train, as_tibble ), 
    test= map(test, as_tibble)
  )
```

To see the different rows in the datasets converted as a dataframe: 
```{r, eval= FALSE}
cv_df|>
  pull(train)|>
  nth(3)|>
  as_tibble()
```

Fit models, extract RMSEs 

- use `map()` to run the linear model on the training dataframes 
- then use `map2_bdl()` to map `rmse()`over model dataframe and test dataframes and return the single number of rmse for that linear model


```{r}

cv_results_df = 
  cv_df |>
  mutate(
    linear_mod = map( train, \(x) lm (logratio ~ range, data = x )), 
    smooth_mod = map( train, \(x) gam(logratio ~ s(range), data = x )), 
    wiggly_mod = map(train, \(x) gam(logratio ~ s(range, k = 30), sp = 10e-6, data= x )
  ))|>
  mutate(
    rmse_linear = map2_dbl( linear_mod, test, rmse ), 
    rmse_smooth = map2_dbl(smooth_mod, test, rmse), 
    rmse_wiggly = map2_dbl(wiggly_mod, test, rmse)
  )

cv_results_df
```

**Cross validation was used to compare datasets and models** 

Looking at the RMSE distirbution: 
```{r}

cv_results_df |>
  select(starts_with("rmse"))|>
  pivot_longer( 
    everything(),
    names_to = "model",
    values_to= "rmse", 
    names_prefix= "rmse_")|> 
  ggplot(aes(x= model, y = rmse))+ 
  geom_violin()
```

The smooth model consistently out predicts the linear model and slighty better than the wiggly model. 
