---
title: "Linear_models"
author: "Kimberly Lopez"
date: "2024-11-07"
output: github_document
---

**Example** 

```{r}
library(tidyverse)
library(p8105.datasets)

set.seed(1)

```

# Model fitting with the Airbnb data 

First some data cleaning 
```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)
```

**Fit a simple model** 

We can get the summary of a fit model using the `summary()` function. We can also use `names()` to know the other variables from the model fitted. 

- we can extract the coefficents of the model using `coef()` 
- or we can use `broom::tidy()` to return values in a tibble
- `broom::glance()`

```{r}
fit = lm(price ~ stars, data = nyc_airbnb)

summary(fit)
names(fit)
coef(fit)

fit|>
  broom::tidy()|>
  select(term,estimate,p.value)
```

**A bit more advance simple model** 

Linear model with two variables 
```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)

fit|> 
  broom::tidy()|>
  select(term,estimate,p.value)|>
  mutate(
    term= str_replace(term, "borough", "Borough: "))|>
      knitr::kable(digits=3)
```

WE reload the data to apply the correct factors in accordance to 

It’s important to note that changing reference categories won’t change “fit” or statistical sigificance, but can affect ease of interpretation.

Now the data will show the same values, but in comparason to the refereance group: 

```{r}
nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))


fit = lm(price ~ stars + borough, data = nyc_airbnb)

fit|> 
  broom::tidy()|>
  select(term,estimate,p.value)|>
  mutate(
    term= str_replace(term, "borough", "Borough: "))|>
      knitr::kable(digits=3)
```

# Tidying output 


# Diagnostics 

**Most diagnostics use residuals** 

First some exploratory data anlysis 
```{r}
 nyc_airbnb|>
  ggplot(aes(x=stars,y=price))+
  geom_point()+
  stat_smooth(method= "lm")
```

The `modelr` package can be used to add residuals and fitted values to a dataframe.
```{r}
modelr::add_residuals(nyc_airbnb, fit)
```


We can plot them with a violin plot 
```{r}
modelr::add_residuals(nyc_airbnb, fit)|>
  ggplot(aes(x=borough, y=resid))+
  geom_violin()+
  ylim(-100,100)
```

We can plot the residua;s against stars which results in a skewed plot. 

```{r}
nyc_airbnb|>
  modelr::add_residuals(fit)|>
  modelr::add_predictions(fit)|>
  ggplot(aes(x=pred, y= resid))+ 
  geom_point()
```

# Hypothesis Testing 

Model summaries include results of t-tests for single coefficients, and are the standard way of assessing statistical significance.

Testing multiple coefficients is somewhat more complicated. A useful approach is to use nested models, meaning that the terms in a simple “null” model are a subset of the terms in a more complex “alternative” model

**When adding more than one variable use `ANOVA`:**

Using a mixed model is reasonable for more than 2 but still low variables.

```{r}
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
```

Anova only works if your doing nested models 

```{r}
fit_null|>
  broom::tidy()

fit_alt|> 
  broom::tidy()

anova(fit_null,fit_alt)|>
  broom::tidy()
```

# Nesting Data

**Do effects differ across broughs**
 
 Does increasing price in manhatten or brooklyn have an affect? Does shared rooms in brooklyn versus manhatten differ?
 
 First use 
```{r}

nyc_airbnb |> 
  lm(price ~ stars * borough + room_type * borough, data = _) |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)
```
 
 Then, We could also fit seperate models to see indivdually how they interact in each borough.
 
 We see effetc of room and stars in Manhattan only 
```{r eval=FALSE}
nyc_airbnb|> 
  filter(borough=="Manhatten")|>
  lm(price ~ stars + room_type, data = _)|>
  broom::tidy()

nyc_airbnb|> 
  filter(borough=="Brooklyn")|>
  lm(price~ stars + room_type, data = _)|>
  broom::tidy()
```
 
 Get fancy and use list columns instead 
 
 We can take data and put it into the argument element in lm which we can then create a new column `model`. We can take it a step further and map again using broom::tidy 

**This is model fitting stuff** 
```{r}

nyc_airbnb|> 
  nest(data = -borough)|>
  mutate(
    model= map(data,\(x) lm(price ~ stars + room_type, data = x) ), 
    reuslts = map(model, broom::tidy)
  )
  
```


We can check specifically for each borough and unset the reuslts column and tidy the data to get values on coefficients for each variable. 

**This is more tidying stuff after model fitting**
```{r}
nyc_airbnb|> 
  nest(data = -borough)|>
  mutate(
    model= map(data,\(x) lm(price ~ stars + room_type, data = x) ), 
    results = map(model, broom::tidy)
  )|>
  select(borough,results)|>
  unnest(results)|>
  select(borough, term, estimate)|>
  pivot_wider ( 
    names_from = term,
    values_from = estimate)
```

**Question: What about room types across manahten neighbohoods?** 

- we are filtering a variable borough to search for a specific borough and neighborhood 

```{r}
nyc_airbnb|> 
  filter( 
    borough== "Manhattan", 
    neighborhood == "Chinatown")|>
  lm(price~ stars+ room_type, data = _ )|>
  broom::tidy()

nyc_airbnb|> 
  filter( 
    borough== "Manhattan", 
    neighborhood == "Chelsea")|>
  lm(price~ stars+ room_type, data = _ )|>
  broom::tidy()
```

The results show they do differ by neighborhodos in the broough. 

We can take this further for all nieghborhoods in Manhattan now which is the same as above but now for all neighborhoods by using `nest` for **borough and neighborhood**

```{r}
manhatten_neighborhood_fit_df = 
  nyc_airbnb|> 
  filter(borough == "Manhattan")|>
  nest(data = -(borough:neighborhood))|>
  mutate(
    model = map (data, \(x) lm(price ~ stars + room_type, data = x )), 
    results = map(model, broom::tidy)
  )|> 
  select(neighborhood, results)|> 
  unnest(results)
```
**Tip: Do things once for a variable, then try to nest if its working, then run the nesting code step by step to make sure no lines of code break, and continu tidying**

We can then look at the effect of room type using the above df but filtering by room_type. 

```{r}
manhatten_neighborhood_fit_df|> 
  filter(str_detect(term, "room_type"))
```

Plot this dataframe: 
```{r}
manhatten_neighborhood_fit_df|> 
  filter(str_detect(term, "room_type"))|> 
  ggplot(aes(x= term, y = estimate))+ 
  geom_boxplot()
```

# Binary Outcomes 

```{r}
baltimore_df = 
  read_csv("data/homicide-data.csv") |> 
  filter(city == "Baltimore") |> 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) |> 
  select(resolved, victim_age, victim_race, victim_sex)

head(baltimore_df)
```


